Inverse conformational selection in lipid-protein binding

### 2021-05-27 manuscriptPGPE.tex
Update title and abstract

\title{Conformational plasticity of lipid headgroups in cellular membranes and protein--lipid complexes} %Title of paper

@@

\begin{abstract}
% insert abstract here
  The chemistry of lipid headgroups, the water-facing interface of cellular membranes, regulates cell functions via lipid--protein interactions. In the physiologically relevant liquid state,
  lipid molecules are not fixed to a single structure but sample an ensemble of conformations with Boltzmann-weighted probabilities.
  However, the distributions of conformations in such ensembles and their lipid-type dependence have not yet been experimentally determined: 
  It is not clear if lipid molecules exchange between a few rigid structures,
  or fluctuate freely across a practically continuous spectrum of  conformations.
  Here, we combine solid-state NMR experiments and molecular dynamics simulations from the NMRlipids Project to resolve the conformational ensembles of headgroups of four key lipid types in various biologically relevant conditions. We find all lipid headgroup types to sample a wide range of conformations in both neutral and charged cellular membranes.
  Differences in the headgroup chemistry do manifest
  in probability distributions of conformations, but all four lipid types can access almost the whole conformation range.
  Together with the analysis of 894 protein-bound lipid structures from the Protein Data Bank (PDB), this suggests that lipids can bind to proteins in a wide range of conformations, which are not limited by the headgroup chemistry. Therefore, the selective adsorption of peripheral proteins to membranes and the binding of lipids to transmembrane proteins are likely to be regulated by specific protein--lipid interactions, not by any inherent conformational restrictions of the lipids.
  Our results pave the way to comprehensive understanding of lipid--protein interactions in biomedical applications, in particular in
  lipid-mediated signaling.

>>>>

\title{Inverse conformational selection in lipid-protein binding} %Title of paper

@@

\begin{abstract}
% insert abstract here
  Interest in lipid interactions with proteins and other biomolecules is emerging not only in
  fundamental biochemistry, but also in the field of nanobiotechnology where lipid nanoparticles
  are commonly used, for example, as carriers in mRNA vaccines.
  The outward facing components of lipid bilayer, the lipid headgroups, regulate how cellular membranes or lipid nanoparticles
  interact with approaching substances, such as proteins, drugs, RNA or viruses.
  %In the physiologically relevant liquid state,
  %lipid molecules are not fixed to a single structure but sample an ensemble of conformations with Boltzmann-weighted probabilities.
  Because conformational ensembles of lipids in physiologically relevant liquid state 
  %However, the distributions of conformations in such ensembles and their lipid-type dependence
  have not yet been experimentally determined, a crucial piece has been missing from the conception on how they interact with other biomolecules: 
  It is not clear if lipid molecules exchange between a few rigid structures,
  or fluctuate freely across a practically continuous spectrum of  conformations.
  Here, we combine solid-state NMR experiments and molecular dynamics simulations from the NMRlipids Project to resolve the conformational ensembles of headgroups of four key lipid types in various biologically relevant conditions. We find that all lipid headgroup types sample a wide range of conformations % a wide range of conformations
  in both neutral and charged cellular membranes, and that differences in the headgroup chemistry manifests only
  in probability distributions of conformations. %, but all four lipid types can access .
  Together with the analysis of 894 protein-bound lipid structures from the Protein Data Bank (PDB), this suggests that lipids can bind to proteins in a wide range of conformations, which are not limited by the headgroup chemistry.
  %Therefore, the selective adsorption of peripheral proteins to membranes and the binding of lipids to transmembrane proteins are likely to be regulated by specific protein--lipid interactions, not by any inherent conformational restrictions of the lipids.
  %Our results pave the way to comprehensive understanding of lipid--protein interactions in biomedical applications, in particular in
  %lipid-mediated signaling.
  Our results propose that lipids are able to select a suitable conformation from the wide range of selection to fit the various binding sites in proteins.
  The proposed {\it inverse conformational selection model} can be also extended to lipid binding to other biomolecules than proteins, such as drug, RNA and viruses.
